---
title: "Section 7: Expressing variability"
format: html
---


# Section 7.2

```{r}
library(tidyverse)
```

```{r}
library(here)
statsurvey_df <- read_csv(here("data/stat113_survey.csv"))
```

The data set has many variables. Some that we will use are:

    time_both, the semester and year the survey was taken
    GPA, the current GPA of the student
    Tattoo, whether or not the student has a Tattoo
    TV, the amount of time spent watching TV per week
    Facebook, the number of Facebook friends

Using this data set, we answer some interesting questions. First, is there evidence of grade inflation at SLU? That is, is there evidence that student GPAs have increased over time? We will start this problem by making a sketch of a couple of visualizations we might want to create.

```{r}
statsurvey_df2 <- statsurvey_df %>% 
  mutate(time_year = as.numeric(time_year)) %>%
  mutate(academic_year = if_else(time_semester == "S", 
                                 true = time_year -  1, 
                                 false = time_year)) %>%
  relocate(academic_year, .after = time_year) %>% 
  filter(GPA <= 4) %>%
  filter(!is.na(GPA)) 


statsurvey_df2 %>%
  group_by(academic_year) %>%
  summarize(mean_gpa = mean(GPA, na.rm = TRUE),
            sd_gpa = sd(GPA),
            n_students = n()) %>%
  mutate(se_gpa = (sd_gpa / sqrt(n_students)),
         lb_se = mean_gpa - se_gpa,
         ub_se = mean_gpa + se_gpa,
         academic_year = academic_year + 2000) %>%
  # now we make our plot using the geom_errorbar(), using the colors, red, white and scarlet
  ggplot(aes(x = academic_year, y = mean_gpa)) +
  geom_errorbar(aes(ymin = lb_se, ymax = ub_se), width = 0.2, color = "darkred") +
  geom_point() +
  geom_line(color = "red") +
  labs(title = "GPA over time",
       x = "Academic year",
       y = "GPA") + theme_minimal()
```



# Section. 7.3

## Exercise 1. 
*Is there evidence from the STAT 113 survey that tattoos have become more or less common (at least among SLU students)? Construct a plot that shows the proportion of students who have a Tattoo in each semester from the STAT 113 survey, along with standard error bars for the estimate in each semester.*

```{r}
statsurvey_df2 %>%
  group_by(time_year) %>%
  # need sort out all of NA's for this column
  filter(!is.na(Tattoo)) %>%
  summarize(n_tattoo = sum(Tattoo == "Yes"),
            n_students = n()) %>%
  mutate(prop_tattoo = n_tattoo / n_students,
         se_tattoo = sqrt(prop_tattoo * (1 - prop_tattoo) / n_students),
         time_year = time_year + 2000) %>%
  ggplot(aes(x = time_year, y = prop_tattoo)) +
  geom_errorbar(aes(ymin = prop_tattoo - se_tattoo, ymax = prop_tattoo + se_tattoo), width = 0.2, color = "darkred") +
  geom_point() +
  geom_line(color = "red") +
  labs(title = "Proportion of Students w/ Tattoos Over Time",
       x = "Academic year",
       y = "Proportion with tattoos") +
  theme_minimal() +
  # lets fix the x-axis to be more readable
  scale_x_continuous(breaks = seq(2011, 2021, 1))
```















