---
title: "Class Prep 5"
author: "Matt Maslow"
format: html
---

# Exercise 1: 
*For each of the core dplyr functions mutate(), arrange(), select(), slice(), filter(), group_by(), and summarise(), write a one sentence summary of the overall purpose of the function.*

**mutate()**- Allows users to manipulate/change/add variables in the data set.

**arrange()**-This will order the imputed argument, or given variable, alpha-numerically, and rest of cols/vars sort themselves accordingly.

**select()**-Allows users to pick which variables they want to view in their final output.

**slice()**-Allows users to split the data at "n" number of rows, creating a separate data set from original.

**filter()**-This allows users to specify what they see in their final output, such as only looking at specific category/level for a particular variable.

**group_by()**-Allows users to categorize the data into groups.

**summarise()**-Produces tibble of imputed information that can be either displayed or stored, and common used to show descriptive statistics of particular variable(s) or data.


# Exercise 2: 
*Review mutate(). Create a new variable in hpi_df that is equal to Wellbeing / Footprint.*
```{r}
library(tidyverse)
hpi_df <- read_csv("data/hpi-tidy.csv")
```
```{r}
hpi_df %>%
  mutate(WellB_FootP = Wellbeing/Footprint) %>%
  relocate(WellB_FootP, Wellbeing, Footprint)
```


# Exercise 3: 
*Review mutate() and case_when(). Create a new variable in hpi_df that is equal to "80s" if LifeExp is in the 80’s, "70s" if LifeExp is in the 70s, and "Below 70" if LifeExp is less than 70.*
```{r}
hpi_df %>%
  mutate(AgeRange = case_when(
    LifeExpectancy >= 80 ~ "80s",
    LifeExpectancy >= 70 ~ "70s",
    LifeExpectancy < 70 ~ "Below 70")) %>%
  relocate(AgeRange, LifeExpectancy)
```


# Exercise 4. 
*Review arrange(). Sort the hpi_df data so that the country with the highest LifeExp is in the first row and the country with the lowest LifeExp is in the last row.*
```{r}
hpi_df %>% arrange(-LifeExpectancy)
```


# Exercise 5. 
*Review select(). Make a data frame from hpi_df that only has the variables Country and LifeExp.*
```{r}
(Country_LifeExp_df <- hpi_df %>%  select(Country, LifeExpectancy)) 
```


# Exercise 6. 
*Review slice(). Make a data frame from hpi_df that only has the first 5 rows.*
```{r}
(First5_hpi <- hpi_df %>% slice(1:5))  %>%pander::pander()
```


# Exercise 7. 
*Review filter(). Make a data frame from hpi_df that only has countries in the "Western World" Region.*
```{r}
(WestWrld_hpi <-hpi_df %>% filter(Region == "Western World"))
```


# Exercise 8. 
*Review filter(). Make a data frame from hpi_df that only has countries with a LifeExp less than 55.*
```{r}
(less55LifExp_hpi <- hpi_df %>% filter(LifeExpectancy < 55))
```


# Exercise 9. 
*Review group_by() and summarise(). Make a table of the number of countries in each Region. (Hint: recall that the n() function is the counting function in dplyr).*
```{r}
hpi_df %>%
  group_by(Region) %>%
  summarise(Country_Count = n()) %>%
  pander::pander()
```


# Exercise 10. 
*Review group_by() and summarise(). Make a table with the maximum LifeExp in each Region.*
```{r}
hpi_df %>%
  group_by(Region) %>%
  summarise(Max_LifeExp = max(LifeExpectancy)) %>%
  pander::pander()
```
















