---
title: "Class Prep 8"
author: "Matt Maslow"
format: html
---

# section 8.1

```{r}
library(tidyverse)
library(broom)
library(here)
theme_set(theme_minimal())

coffee_df <- read_csv(here("data/coffee_ratings.csv"))
coffee_mod <- lm(total_cup_points ~ species + aroma + flavor +
                   sweetness + moisture,
   data = coffee_df)
```

```{r}
summary(coffee_mod)
```
```{r}
summary(coffee_mod)$coefficients["moisture", 4]
```


```{r}
tidy(coffee_mod)
```


```{r}
tidy(coffee_mod) |> select(p.value)
```

```{r}
tidy(coffee_mod) |> filter(term == "aroma")
```

## Section 8.1.2: glance()

```{r}
glance(coffee_mod)
```

## Section 8.1.3: augment()

```{r}
augment(coffee_mod)
```

```{r}
augment_df <- augment(coffee_mod)
augment_df |> filter(.cooksd > 1)
```

```{r}
ggplot(data = coffee_df, aes(x = total_cup_points)) +
  geom_histogram(bins = 15, fill = "white", colour = "black")
```
```{r}
augment_df |> filter(.hat > 0.2)
```

```{r}
augment_df |> filter(.std.resid > 3 | .std.resid < -3)
```

```{r}
ggplot(data = augment_df |> filter(.fitted > 25),
       aes(x = .fitted, y = .resid)) +
  geom_point() 
```

## Exercises

### Exercise 1. 
*Examine the penguins data set in the palmerpenguins package:*

```{r}
library(palmerpenguins)
penguins
```

*Fit a linear regression model with body_mass_g as the response variable and species and bill_length_mm as the predictors. Note that penguins with missing values for any of these three variables will be dropped from the analysis.*

```{r}
penguins_mod <- lm(body_mass_g ~ species + bill_length_mm, data = penguins)
summary(penguins_mod)
```

### Exercise 2. 
*Create a table of summary output, including coefficient estimates, standard errors, test statistics, and p-values, using one of the broom functions.*

```{r}
tidy(penguins_mod)
```


### Exercise 3. 
*Use glance() to glance at some of the relevant model statistics.*

```{r}
glance(penguins_mod)
```


### Exercise 4. 
*Using augment(), create a plot of the residuals vs. the fitted values and evaluate the constant variance assumption.*

```{r}
augment(penguins_mod) |> ggplot(aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

Look at the fits versus residuals plot, can conclude that the residuals are roughly lineaer at zero, but not entirely.


### Exercise 5. 
*Using augment(), check to see if there are any penguins that are influential. Use 0.75 as your cut-off value for Cook’s distance.*

```{r}
augment(penguins_mod) |> filter(.cooksd > 0.75)
```

No influential penguins are present in this model

















